ForestClaw: Hybrid forest-of-octrees AMR for hyperbolic conservation laws

### 2013-02-26 docs/parco13/abstract.tex
Write intro

Experience from integration with large-scale adaptive-mesh PDE solvers.

Design goals: Lightweight, modular, reuse.

@@

p4est \cite{BursteddeWilcoxGhattas11}

rank as is common with parallelization using the MPI framework
\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.

@@

Iterators over all leafs, optionally restricted to a given level.

@@

neighbor processor: data needs to be fetched over the network first.

@@

@@

%%%%%%%%%%% The bibliography starts:
\bibliographystyle{siam}
\bibliography{ccgo}

>>>>

With the advent of high-throughput coprocessors, such as GPGPUs or the MIC
chip, comes the opportunity to sustain unprecedented rates of floating point
operations at comparably high integration density and low cost.  These
architectures, however, require careful structuring of the data layout and
memory access patterns to fully utilize their multithreading and vectorization
capabilities.

Consequently, it is not clear a priori how to accelerate PDE solvers that use
adaptive mesh refinement, especially when working with unstructured meshes.  It
has been realized early though that it helps to aggregate degrees of freedom
(DOF) at the element level, as it is done with high-order spectral element
\cite{TufoFischer99}, finite volume, or discontinuous Galerkin
\cite{BursteddeGhattasGurnisEtAl10} methods.
%
% For example, SEs optimized for computational speed have originally been
% implemented on unstructured conforming meshes \cite{TufoFischer99} and later
% been extended to non-conforming adaptive meshes \cite{FischerKruseLoth02,
% BursteddeGhattasGurnisEtAl10}.

To enable hardware acceleration for dynamic AMR, we like to build upon the
forest-of-octrees paradigm because of its low overhead and parallel scalability
\cite{BursteddeWilcoxGhattas11}.  We go beyond the traditional high-order
element by introducing a complete compute patch with $m^d$ DOFs as our
macro-element.  In fact, this approach resembles block-structured AMR
\cite{ColellaGravesKeenEtAl07}
% todo cite John B.\ Bell
except that the patches are not overlapping, which enables us to capitalize on
our previous experience with AMR PDE solvers
\cite{BursteddeStadlerAlisicEtAl13}.

\section*{Design principles}

Now combine.

Experience from integration with large-scale adaptive-mesh PDE solvers.
Design goals: Lightweight, modular, reuse.

@@

rank as is common with parallelization using the MPI framework
%\cite{Forum94, SnirOttoHuss-LedermanEtAl96}.

@@

$z$-order

Iterators over all leafs, optionally restricted to a given level.
Random access is possible.  Looping over the patches in the order prescribed
by the forest leads to a high percentage of cache reuse, since the ghost
patches are likely closeby with respect to the $z$-order.

@@

neighbor rank: data needs to be fetched over the network first.

@@

Patch: contains the local coordinates and copies of neighboring ghost degrees
of freedom.  The layer of ghost copies is usually 2 points deep.

@@

%%%%%%%%%%% The bibliography starts:
\bibliographystyle{unsrt}
\bibliography{ccgo}
